The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-05-26 paper/referee-reply.tex

\title{Response to the referee}

\begin{document}

\date{}
\maketitle

\section{Global Comments}

\subsection{Variation with the viewing angle}

The problematic of anisotropy is only studied from Sect3.4, the former sections consider global quantities (spectral shapes, escape fractions, etc...). This is ok only if there is NO anisotropy induced by rotation, which is not obvious, a priori. You should make it clear from the beginning, telling that you will investigate anisotropy at the end of the paper only, because you checked that lya properties are isotropic, at least in the range of parameters that you investigated so far.\\

\textbf{ We have decide to study in this paper the case in which there is No anisotropy. The part concerning the off-center emission is going to be presented in a later study\\}

As explained in sect 3.4, rotation kills the spherical symmetry of your problem, and the rotation axis defines a preferential direction. We could expect some variation of the emerging flux, the lyman-alpha escape fraction, and the spectral shape, with viewing angle. From Fig8, it seems that the flux is the same in all directions for spherical distributions of sources. This important result is not enough emphasized.\\

\textbf{We do not find the flux variation with the viewing angle and now as highlighted in the new text we emphasize more this new result, also we made a new plot Figure 5 in the paper, showing this results for both distributions (homogeneous/central) and for optical depth $\tau_{H}=10^{5}$ we choose these models because this are the ones 
that are more sensitive to rotation.\\}

Do you see any difference in the spectral shape with viewing angle ? To illustrate this point, it would be interesting to build a 2D plot as you did on Fig 5, but replace Nb scat in ordinates by the viewing angle μ. We could immediatly see if there is an evolution of the shape with viewing angle or not. If you find NO evolution of the Lya shape with viewing angle, I think that it is an interesting counter-intuitive result, that you should advertise more.\\

\textbf{Thanks very much for this recommendation we made this new plot and we do find variations in the morphology of the line due to variations in the viewing angle. We have restructure the paper this new plots now correspond to figure 1 & 2 in the paper.  Also we emphasize this results in the results, discussion and conclusion sections.\\}

What about the variation of lya escape fraction with viewing angle ? I propose that in the beginning of Sect.3 Results, you first emphasize that, maybe counter-intuitively, you did not find any variation of the Lya properties with viewing angle, so you will present first angle average lya properties, and you will come back to the anisotropy problematic only at the end on the paper.\\

\textbf{As escape fraction depends on the number of scatterings and this is invariant for different viewing angles the escape fraction does not change with the viewing angle. (MAKE A PLOT AND EXPLAIN THIS BETTER)
}

@@

\subsection*{Fig.3}

\subsection*{Fig4 and 6}

Waiting for the cluster in germany!

\subsection*{Fig5}
\begin{figure}
\includegraphics[scale=0.4]{Histogram2dNscattVSRadius.png}
\end{figure}

### 2014-05-26 paper/referee-reply.tex

As explained in sect 3.4, rotation kills the spherical symmetry of your problem, and the rotation axis defines a preferential direction. We could expect some variation of the emerging flux, the lyman-alpha escape fraction, and the spectral shape, with viewing angle. From Fig8, it seems that the flux is the same in all directions for spherical distributions of sources. This important result is not enough emphasized.\\

\textbf{We do not find the flux variation with the viewing angle and now as highlighted in the new text we emphasize more this new result, also we made a new plot Figure 5 in the paper, showing this results for both distributions (homogeneous/central) and for optical depth $\tau_{H}=10^{5}$ we choose these models because this are the ones 
that are more sensitive to rotation.\\}

Do you see any difference in the spectral shape with viewing angle ? To illustrate this point, it would be interesting to build a 2D plot as you did on Fig 5, but replace Nb scat in ordinates by the viewing angle μ. We could immediatly see if there is an evolution of the shape with viewing angle or not. If you find NO evolution of the Lya shape with viewing angle, I think that it is an interesting counter-intuitive result, that you should advertise more.\\

\textbf{Thanks very much for this recommendation we made this new plot and we do find variations in the morphology of the line due to variations in the viewing angle. We have restructure the paper this new plots now correspond to figure 2 & 3 in the paper.  Also we emphasize this results in the results, discussion and conclusion sections.\\}

@@

We want to study the pure effect of rotation on the spectra, off course it would
be interesting to study the effect of rotation on systems with outflows, this paper
is under preparation by M.C Remolina-Gutierrez et al.

\section*{Details}

\subsection*{Introduction}

With Orsi et al 2012, please cite also Garel et al 2012. With Zheng&Wallace 2013, please cite also Behrens et al 2014.\\

\textit{This changes were performed.}
\subsection*{Fig.1}

I guess that the spectra presented in Fig1 are integrated over all directions, right ? You should describe explicitly how you build them. You could skip the x notation in absciss, as it is not used in the discussion, whereas the velocity is used to compare to FWHM, on Fig2.\\

\textbf{Now this figure is the Fig.4 in the paper. what we did was to fix a viewing angle in this case $\theta= \pi/2$ and make the histogram of the photons frequency.}

@@

\begin{itemize}
\item Description about how we build Fig.1 (Done)
\item Fig 1. remove x notation (Done)
\end{itemize}

\subsection*{Fig.2}

Fig.2 Can you explain how you measure the FWHM of a double-peaked profile ? Do you fit it with a gaussian ?

We don't fit a gaussian, we interpolate points to the line profile and then  we measure the width by finding
the closest points to the half maximum in our histogram.

### 2014-05-27 paper/referee-reply.tex

\title{Response to the referee}

\begin{document}

\date{}
\maketitle

Dear referee the main changes in our paper are that we do find

@@

The problematic of anisotropy is only studied from Sect3.4, the former sections consider global quantities (spectral shapes, escape fractions, etc...). This is ok only if there is NO anisotropy induced by rotation, which is not obvious, a priori. You should make it clear from the beginning, telling that you will investigate anisotropy at the end of the paper only, because you checked that lya properties are isotropic, at least in the range of parameters that you investigated so far.\\

\textbf{ We have decide to study in this paper the case in which there is No anisotropy. The part concerning the off-center emission is going to be presented in a later study\\}

As explained in sect 3.4, rotation kills the spherical symmetry of your problem, and the rotation axis defines a preferential direction. We could expect some variation of the emerging flux, the lyman-alpha escape fraction, and the spectral shape, with viewing angle. From Fig8, it seems that the flux is the same in all directions for spherical distributions of sources. This important result is not enough emphasized.\\

\textbf{We do not find the flux variation with the viewing angle and now as highlighted in the new text we emphasize more this new result, also we made a new plot Figure 5 in the paper, showing this results for both distributions (homogeneous/central) and for optical depth $\tau_{H}=10^{5}$ we choose these models because this are the ones 
that are more sensitive to rotation.}\\

Do you see any difference in the spectral shape with viewing angle? To illustrate this point, it would be interesting to build a 2D plot as you did on Fig 5, but replace Nb scat in ordinates by the viewing angle $\mu$. We could immediatly see if there is an evolution of the shape with viewing angle or not. If you find NO evolution of the Lya shape with viewing angle, I think that it is an interesting counter-intuitive result, that you should advertise more.\\

\textbf{Thanks very much for this recommendation we made this new plot and we do find variations in the morphology of the line due to variations in the viewing angle. We have restructure the paper this new plots now correspond to figure 2 \& 3 in the paper.  Also we emphasize this results in the results, discussion and conclusion sections.\\}

@@

We want to study the pure effect of rotation on the spectra, off course it would
be interesting to study the effect of rotation on systems with outflows, this paper
is under preparation by M.C Remolina-Gutierrez et al.

\section*{Details}

\subsection*{Introduction}

With Orsi et al 2012, please cite also Garel et al 2012. With Zheng \& Wallace 2013, please cite also Behrens et al 2014.\\

@@

Fig.2 Can you explain how you measure the FWHM of a double-peaked profile ? Do you fit it with a gaussian ?

We don't fit a gaussian, we interpolate points to the line profile and then  we measure the width by finding
the closest points to the half maximum in our histogram.

\subsection*{Fig.3}

Do you have an idea why the (central source, intermediate optical depth) case with Vmax=300 has a single plateau instead of 2 peaks ? Do you find this with the two codes ?

@@

\subsection*{Fig4 and 6}

This is a surprising result that the number of scatterings (escape fraction) stays constant as the rotation velocity increases, for a central source, whereas the global spectral shape is altered. Did you try with higher/extreme values of Vmax (=1000 km/s, even if not physically motivated) ? Do you believe that the number of scatterings decreases with very high values, or that it is independant of the rotation velocity ? Is the escape fraction from a dusty rotating cloud with central source independant of the rotation velocity ?

\subsection*{Fig5}
This is a very nice figure ! Looking at the top right panel, with its “photosphere”, I’m surprised that the distribution of Nscatt is bipolar, I would have expected a smooth transition between the 2 regimes. Do you have a physical explanation why photons escape after either (less than 10) or (more than 1000) scatterings, and none escape with 100 scatt ? To test the ’photosphere’ assumption, you could also do the radiation transfer in a cloud with the distribution of sources only in an inner sphere, and another experienment with sources only in an outer shell. If I understood corectly, you would expect thant the bottom spot on the top right panel is made of photons from the outer shell, and the uppper spot from photons emitted inside the cloud.

\begin{figure}
\includegraphics[scale=0.4]{Histogram2dNscattVSRadius.png}
\end{figure}

\subsection*{End of paragraph 3.3}

One sentence is not clear : we see that the escape fraction does not increase significantly from $\tau$ = 105 to $\tau = 10^{6}$. This counter-intuitive result.... It sounds like you were expecting a strong increase... A decrease is expected from $\tau$ = 105 to $\tau$ = 106, not an increase, but indeed on the graphe we can see an unexpected (small) increase. I do not understand the explanation for this behaviour.
(Mark)

\subsection*{Fig8}

Referred to as Figure 7 in the text (paragraph 3.4). To my mind, this figure illustrates the main result of your study, it has to be more demonstrative. On this figure, the numerical noise seems very big compare to the small number of bins, bigger than on the spectra from Fig1, how is it possible ? Could you re-do this figure with more photons, maybe with more bins ? Could you plot an histogram instead of broken lines, to ease the reading ? In the text, you write anisotropy induced by rotation is at the 3$\%$ level, how do to estimate this number? I think that the noise is at $3\%$ level, and the distribution is compatible with ’flat’. In the corresponding text, you first say that $F(\mu)$ does not depend on Vmax. Again, if real, this is an important result, that you must advertise more. Please show it on this Fig, by comparing F(μ) without rotation (in red, for example) to the other distributions that you get with different values of Vmax. Then, you mention that for high optical depth values, $F(\mu)$ (not the variations of $\mu$) can vary up to $15\%$. First, what do you call high optical depth ? your highest setup is τ = 107, not extremely high for lyman alpha ”standards’, corresponding to NHI ∼ 1020 cm−2. So, if you see a trend with optical depth, I encourage you to check higher optical depth regimes, corresponding better to galaxy scales. If you can probe that anisotropy induced by rotation is rising with optical depth, this would be an interesting result.
If higher optical depth regimes lead to anisotropic escape, could you check the effet of anisotropy on the lya spectral shape ?

### 2014-05-30 paper/ms.tex
revised abstract, conclusions and figure captions

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$
emission line in distant galaxies. We model a galaxy as a sphere with
an homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We consider cases 
of a single central \lya source, and a uniform distribution of sources. 
Our main result is that rotation introduces a dependence of the
line morphology with  viewing angle. Observations parallel to the
rotation axis yield lines similar to the static case. The greatest
difference is observed perpendicular to the rotation axis where the
intensity at the line center increases with rotational velocity.  For
homogeneously distributed sources the line becomes single peaked  at
rotational velocities larger than half the line width in the static
case. However, rotation does not not induce any significant anisotropy
in the total line intensity or in the escape fraction. 
\end{abstract}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f1.pdf}
\end{center}
\caption{Geometry of the gas distribution. The angular velocity vector
  is parallel to the unit vector $\hat{k}$. In order to describe the
  departures from spherical symmetry we use the polar angle $\theta$
  formed by the direction of the outgoing photons with respecto to the
  $z$-axis. We define define the variable $\mu\equiv\cos\theta$ to
  report to present our results.
    \label{fig:geometry}}  
\end{figure}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f2.pdf}
\end{center}
\caption{
2D histogram showing the number of photons that escape with frecuency
$x$ forming an angle $\theta$ (parameterized as $\cos\theta$) with the
rotationa axis. The rotational velocity ($0,100,200,300$\kms)
increases from left to right and the optical depth ($10^5$, $10^6$,
$10^7$) from top to bottom. The \ly photons are initialized at the
center of the sphere.
\label{fig:CentralSpec} }   
\end{figure*}

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f4.pdf}
\end{center}
\caption{Shape of the \ly line for different maximum rotational
  velocities for a viewing angle perpendicular to the rotation axis
  ($|\mu|\sim 0$). The continuous (dashed) line represents the central
  (homogeneous) source distributions.  The panels follow the same
  correspondence as Figures \ref{fig:CentralSpec} and \ref{fig:HomSpec}.
    \label{fig:differentvelocities}}  
\end{figure*}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f5.pdf}
\end{center}
\caption{Integrated flux distribution, $F(\mu)$, as a function of the
  viewing angle as parameterized by $\mu$. We show the results for the
  homogeneous source distribution in models with rotational velocities
  $0$\kms, $300$\kms and  optical depths $10^5$, $10^6$ and $10^7$. We
  see that the distribution is flat, meaning that the flux for all
  viewing angles is the same. The results for the central sources are
  similar. 
\label{fig:muhisto}}
\end{figure}

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f6.pdf}
\end{center}
  \caption{FWHM for the non-dusty models as a function of the viewing
  angle parameterized by $\mu\equiv\cos\theta$. 
  Continuous (dashed) lines  correspond to central (homogeneous)
  source distributions. The general trend is of an decreasing FWHM as
  the line of sight becomes parallel to the rotation axis. 
  \label{fig:widthvsmu}} 
\end{figure*}

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=0.45\textwidth]{nscattvsmu.eps}
    \includegraphics[width=0.45\textwidth]{f4.eps}
\end{center}
\caption{Logarithm of the average number of scatterings as function of
  $\mu$ (left panel) and the maximum rotational velocity $V_{\rm
    max}$. The left panel shows the behaviour for $\tau=10^{5}$ and
  $V_{max}=300$\kms but this independence of the independence between
  $N_{\rm scatt}$ and $\mu$ is kept for all models. 
  In the right panel the continuous (dashed) lines represent an
  central (homogeneous) distribution of sources; this shows the clear
  decrease in the number of scatterings for the homogenously
  distributed sources.
\label{fig:Nscatt} }    
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.48\textwidth]{f5_1.eps}
  \includegraphics[width=0.48\textwidth]{f5_2.eps}
  \includegraphics[width=0.48\textwidth]{f5_3.eps}
  \includegraphics[width=0.48\textwidth]{f5_4.eps}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left panels 
      corresponds to the static case and right panels correspond to
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}}   
\end{figure*}

@@

This paper quantifies for the first time in the literature the effects
of gas rotation in the morphology of the \ly emission line in star forming
galaxies.  The results are based on the study of an homogeneous sphere
of gas with solid body rotation. We explore a range of models by varying
the rotational speed, hydrogen optical depth, dust optical depth and
initial distribution of \ly photons with respect to the gas
density. As a cross-validation, we obtain our results from two
independently developed Monte-Carlo radiative transfer codes.

Our main result is two-fold. First,rotation clearly impacts the \ly line
morphology. Second, rotation introduces an anisotropy for different
viewing angles. For viewing angles close to the poles the line is
double peaked and it makes a transition to a single peaked line for
high rotational velocities and viewing angles along the equator. This
trend is stronger for spheres with homogeneusly distributed radiation
sources.  In contrast, the total flux in the line remains independent
of the viewing angle.

The escape fraction remains constant for diffenret viewing angles only
in the case of centrally distributed sources. In contrast, for an homogeneous
distribution the escape fraction decreases with high rotation 
fraction remain constant for different viewing angles. We showed that
these results can be interpreted in terms of the average number of
scatterings for the photons, a quantity that also remains constant
independent of the viewing angle and the rotational velocity for
centrally distributed sources and decreases for homogenously
distributed source.

Quantitatively the main results of our study are summarized as
follows.

@@

\item Rotation induces a considerable asymmetry in the morphology 
 of the spectra when the viewing angle $\theta$. This occurs for
 all of our models. The asymmetry is such that observeds with
 $\cos\theta\equiv \mu\sim 1$, i.e. observes aligned with the rotation
 axis, the line is double peaked while for $\mu\sim 0$ and high
 rotation veolcities the linebecomes double peaked.

\item For $\mu \sim 0$ The line width increases with rotational velocity. The line width
  approximately follows the functional for  ${\rm FWHM}^2 = {\rm FWHM}_{
    0}^2 + (V_{\rm max}/\lambda)^2$ where FWHM$_{0}$ indicates the line
  width for the static case and $\lambda$ is a constant determined from
  the radiative transfer results it is $\lambda_{\rm c}=0.84 \pm 0.06$ and
  $\lambda_{\rm h}=0.54\pm 0.10$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homogeneously distributed sources. These cases occur when
  the rotational velocity is larger than half FWHM$_0$ and $\mu\sim 1$.

\item 
For central sources we find that the escape fraction does not change
with rotational velocity nor with the viewing angle. We interpret this
in terms of the average number of scatterings which does not
significantly changes with rotation. This is due to the  
  fact that the vast majority of scatterings events are resonant, 
  during which the mean free path is very short and the effect of gas
  bulk rotation is not enough to affect \ly photons.

@@

@@

>>>>

We present results of radiative transfer calculations to measure the
impact of gas bulk rotation on the morphology of the Lyman $\alpha$ 
emission line in distant galaxies. We model a galaxy as a sphere with
an homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with maximum
velocities in the range $0-300$ \kms and neutral hydrogen optical
depths in the range $\tau_{\rm H}=10^{5}-10^{7}$. We consider cases 
of a single central \lya source, and a uniform distribution of sources. 
Our main result is that rotation introduces a dependence of the
line morphology with  viewing angle. Observations with a line of sight
parallel to the rotation axis yield line morphologies similar to the
static case. The greatest difference is observed for lines of sight
perpendicular to the rotation axis. In this case the intensity at the line
center increases and the line width increase with rotational velocity.  For
homogeneously distributed sources the line becomes single peaked  at
rotational velocities larger than half the line width in the static
case. However, rotation does not not induce any significant anisotropy
in the total line intensity nor in the escape fraction. 
\end{abstract}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f1.pdf}
\end{center}
\caption{Geometry of the gas distribution. The angular velocity vector
  is parallel to the unit vector $\hat{k}$. In order to describe the
  departures from spherical symmetry we use the polar angle $\theta$
  formed by the direction of the outgoing photons with respecto to the
  $z$-axis. We define define the variable $\mu\equiv\cos\theta$ to
  report to present our results. Computing the spectra for photons in
  a   narrow range of $\mu$ is equivalent to having a line-of-sight
  oriented in that direction.  
    \label{fig:geometry}}  
\end{figure}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f2.pdf}
\end{center}
\caption{
2D histogram showing the number of photons that escape with frecuency
$x$ forming an angle $\theta$ (parameterized as $|\cos\theta|$) with the
rotation axis. The rotational velocity ($0,100,200,300$\kms)
increases from left to right and the optical depth ($10^5$, $10^6$,
$10^7$) from top to bottom. The \ly photons are initialized at the
center of the sphere.
\label{fig:CentralSpec} }   
\end{figure*}

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f4.pdf}
\end{center}
\caption{Shape of the \ly line for different maximum rotational
  velocities for a LoS perpendicular to the rotation axis
  ($|\mu|\sim 0$). The continuous (dashed) line represents the central
  (homogeneous) source distributions.  The panels follow the same
  distribution as in Figures \ref{fig:CentralSpec} and \ref{fig:HomSpec}.
    \label{fig:differentvelocities}}  
\end{figure*}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f5.pdf}
\end{center}
\caption{Integrated flux distribution as a function of the
  viewing angle as parameterized by $\mu$. We show the results for the
  homogeneous source distribution in models with rotational velocities
  $0$\kms, $300$\kms and  optical depths $10^5$, $10^6$ and $10^7$.
  The distributions are flat, meaning that the integrated flux for all
  viewing angles is the same. The results for the central sources are
  similar. 
\label{fig:muhisto}}
\end{figure}

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.95\textwidth]{f6.pdf}
\end{center}
  \caption{FWHM for the non-dusty models as a function of the viewing
  angle parameterized by $|\cos\theta|$. Continuous (dashed) lines  correspond
  to central (homogeneous) source distributions. The general trend is
  of an decreasing line width as the line of sight becomes parallel to the
  rotation axis.
  \label{fig:widthvsmu}} 
\end{figure*}

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=0.45\textwidth]{nscattvsmu.eps}
    \includegraphics[width=0.45\textwidth]{f4.eps}
\end{center}
\caption{Logarithm of the average number of scatterings as function of
  $\mu$ (left panel) and the maximum rotational velocity $V_{\rm
    max}$. The left panel shows the behaviour for $\tau=10^{5}$ and
  $V_{max}=300$\kms as a function of $|\cos\theta|$; the independence of
  $N_{\rm scatt}$ with $\mu$ is present in all models. 
  In the right panel the continuous (dashed) lines represent an
  central (homogeneous) distribution of sources; there is a clear
  decrease in the number of scatterings for the homogenously
  distributed sources. 
\label{fig:Nscatt} }    
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.48\textwidth]{f5_1.eps}
  \includegraphics[width=0.48\textwidth]{f5_2.eps}
  \includegraphics[width=0.48\textwidth]{f5_3.eps}
  \includegraphics[width=0.48\textwidth]{f5_4.eps}    
\end{center}
    \caption{2D histogram of $N_{\rm scatt}$ vs $x$. The upper (lower) panels
      show the homogeneous (central) source distribution. Left panels 
      corresponds to the static case and right panels correspond to
      $V_{\rm max}=300\kms$. The color scale is logarithmic on the
      number of photons with given values of $N_{\rm scatt}$ and
      $x$. \label{fig:Nscatt2D}. A results that stands out is in the upper
      right panel, showing that in homogeneously distributed sources a
      large number of photons can escape escape after a few
      scatterings, $N_{\rm scatt}<10$.}   
\end{figure*}

@@

In this paper we quantified for the first time in the literature the effects
of gas bulk rotation in the morphology of the \ly emission line in star forming
galaxies.  Our results are based on the study of an homogeneous sphere
of gas with solid body rotation. We explore a range of models by varying
the rotational speed, hydrogen optical depth, dust optical depth and
initial distribution of \ly photons with respect to the gas
density. As a cross-validation, we obtained our results from two
independently developed Monte-Carlo radiative transfer codes.

Two conclusions stand out from our study. First, rotation clearly impacts the \ly line
morphology; the width and the relative intensity of the center of the
line and its peaks are affected. Second, rotation introduces an
anisotropy for different viewing angles. For viewing angles close to
the poles the line is double peaked and it makes a transition to a
single peaked line for high rotational velocities and viewing angles
along the equator. This trend is clearer for spheres with
homogeneously distributed radiation sources than it is for central
sources.

In contrast, we found two quantities that are invariant with respect
to the viewing angle and the rotational velocity: the integrated flux and the escape
fraction.  The exception are the models with homogeneously distributed
sources, which present an escape fraction dependent on the rotational
velocity, but still independent of the viewing angle. We showed that
these results could be interpreted in terms of the average number of
scatterings for the photons.

Quantitatively, the main results of our study are summarized as
follows.

@@

\item In all of our models, rotation induces changes in the line morphology
 for different values of the angle between the rotation
 axis and the LoS, $\theta$.  The changes are such that for 
 $\cos\theta\equiv \mu\sim 0$, i.e. LoS perpendicular to the
 rotation axis, and high rotational velocities the line becomes single peaked.

\item For a viewing angle with $\mu=0$ the line width increases with rotational
  velocity. This change approximately follows the functional form  ${\rm FWHM}^2 = {\rm FWHM}_{ 0}^2 + (V_{\rm max}/\lambda)^2$, where FWHM$_{0}$ indicates the line
  width for the static case and $\lambda$ is a constant. We have
  determined this constant to be  $\lambda_{\rm c}=0.84 \pm 0.06$ and
  $\lambda_{\rm h}=0.54\pm 0.10$ for the central and homogeneous source
  distributions, respectively.

\item At fixed rotational velocity the line width decreases as $|\mu|$
  increases, i.e. the smallest value of the line width is observed for
  a line of sight parallel to the ration axis. However the lowest
  values are always larger than the values for the static case.

\item The single peaked line emerges for homogeneously distributed sources 
  when the rotational velocity is larger than half FWHM$_0$ and $\mu\sim
  1$.

\item  For central sources we find that the escape fraction does not change
  with rotational velocity nor with the viewing angle. We interpret this
  in terms of the average number of scatterings which does not
  significantly change with rotation. This is due to the  
  fact that the vast majority of scatterings events are resonant, 
  during which the mean free path is very short and the effect of gas
  bulk rotation is not enough to affect \ly photons.

@@

The definitive and clear impact of rotation on the \ly morphology
suggests that this is an effect that should be taken into accunt at
the moment of interpreting high resolution spectroscopic data. To this
end, we defer a discussion of observed lines features in the context of gas
bulk rotation to a future publication (Remolina-Gutierrez et al., in
preparation).

@@

We thank the referee for the suggestions that allowed us to greatly
improve and better frame the interpretation of our simulations.
